---
title: "01_wiz"
author: "Julian Morales"
date: "2026-10-01"
output: html_document
---
```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

Now we have everything we need

```{r}
weather_df
```

Let's make a scatterplot
```{r}
ggplot(weather_df, aes(x=tmin, y= tmax))+
  geom_point()
```

Preferred to alwasy do the dataframe first

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()
```

Or make it a separate entitiy for saving

```{r}
ggp_temp_scatterplot = 
  weather_df |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()

ggp_temp_scatterplot
```

Make it fancier...
```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color = name))+ #Aesthetic mappings: color by variable name
  geom_point(alpha = .25)+ #alpha makes points a bit transparent
  geom_smooth(se = FALSE) #If you want to see a trend clearly
```

Aesthetics are up to you

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color = name), alpha = .25)+ #you can define aesthetics wherever you need it
  geom_smooth(se = FALSE)
```

No need to HAVE to HAVE a scatterplot

```{r}
# Just gives smooth line not a scatterplot
weather_df |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_smooth(se = FALSE)
```

Show faceting
```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha = .5)+
  facet_grid(.~name) #Separates scatterplot by name 
```
what is done now
```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point(alpha = .5)+
  facet_grid(cols=vars(name)) #does same thing as above
```


Let's look at something else

```{r}
weather_df |> 
  ggplot(aes(x=date, y=tmax, color = name))+
  geom_point(aes(size = prcp), alpha = .5)+
  geom_smooth(se =FALSE)+
  facet_grid(.~name)
```

Make a plot of central park tmx v tmin only, and convert temperatures to fahrenheit.

```{r}
# This has not permanently changed the dataframe, just has done it within this call.
weather_df |> 
  filter(name =="CentralPark_NY") |> 
  mutate(
    tmax = tmax*(9/5) + 32,
    tmin = tmin*(9/5) + 32
  ) |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()
```


what's a hex plot

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_hex() #makes hexagons for plots
```

## Univariate Plots

```{r}
weather_df |> 
  ggplot(aes(x=tmax))+
  geom_histogram()
```


```{r}
weather_df |> 
  ggplot(aes(x=tmax, fill = name))+
  geom_histogram(position = "dodge")
```
Might look better than overlap but has it's own issues
```{r}
weather_df |> 
  ggplot(aes(x=tmax, fill = name))+
  geom_histogram()+
  facet_grid(.~name)
```


Density plots are great!!!

```{r}
weather_df |> 
  ggplot(aes(x = tmax, color = name))+
  geom_density()

#OR
weather_df |> 
  ggplot(aes(x = tmax, fill = name))+
  geom_density(alpha = .3)
```

boxplots
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax))+
  geom_boxplot()
```

violin
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax))+
  geom_violin()
```


Ridge plots...

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = name))+
  geom_density_ridges()
```

##Save some of the plots 

```{r}
ggplot_weather =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name))+
  geom_point(aes(size = prcp), alpha = .5)+
  facet_grid(.~name)

ggplot_weather

ggsave("images/ggplot_weather.pdf", ggplot_weather)
```

```{r fig.width=12, fig.height=12}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()
```

