---
title: "02_viz"
author: "Julian Morales"
date: "2026-10-06"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

Colors!!!

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point()
  scale_color_hue(h = c(100, 100))
```

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point()+
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )+
  theme(legend.position = "bottom")
```

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax, color = name))+
  geom_point()+
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )+
  theme_classic()+
  theme(legend.position = "bottom")
```

update tmax vs date plot

```{r}
weather_df |> 
  ggplot(aes(x=date, y=tmax, color = name))+
  geom_point()+
  geom_smooth(se = FALSE)+
  labs(
    title = "Seasonal trends in Max Temp",
    x = "date",
    y = "Max temp",
    caption = "Max daily temp in three weather stations in 2021 and 2020",
    color = "Location"
  )+
  viridis::scale_color_viridis(
    discrete = TRUE
  )+
  theme_minimal()+
  theme(legend.position = "bottom")
```

## Two More Weird Things

```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(molokai_df, aes(x = date, y = tmax, color = name))+
  geom_point()+
  geom_line(data = central_park_df)
```



Multiple panels with different plot types stiched together by patchwork

```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name))+
  geom_point()

ggp_prcp_density = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name))+
  geom_density(alpha = .5)

ggp_seasonal = 
  weather_df |> 
  ggplot(aes(x=date, y=tmax, color = name))+
  geom_point()
```

we now have everything we need! 

```{r}
library(patchwork)

ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name))+
  geom_point()+
  theme(legend.position = "none")

ggp_prcp_density = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name))+
  geom_density(alpha = .5)+
  theme(legend.position = "none")

ggp_seasonal = 
  weather_df |> 
  ggplot(aes(x=date, y=tmax, color = name))+
  geom_point()+
  theme(legend.position = "none")

(ggp_tmax_tmin - ggp_prcp_density)/ggp_seasonal
```

## Data manipulation

Start with factors

box plots

```{r}
weather_df |>  
  mutate(name = factor(name)) |> 
  ggplot(aes(x = name, y = tmax))+
  geom_boxplot()
```

```{r}
weather_df |>  
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax))+
  geom_boxplot()
```

```{r}
weather_df |> 
  mutate(name = fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax))+
  geom_boxplot()
```


Make practice distribution plot

rather than having tmax and tmin as separate columns, combine data through pivot longer to store if its max/min in one column and the actual temp values in another.
```{r}
weather_df |> 
  select(name, tmax, tmin) |> 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temp"
  ) |> 
  ggplot(aes(x = temp, fill = observation))+
  geom_density(alpha = .5)+
  facet_grid(.~name)
```

Another example

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv") |> 
  janitor::clean_names()

litter_df = 
  read_csv("data/FAS_litter.csv") |> 
  janitor::clean_names()

##FINISH EXAMPLE
```

